---
title: "Hakalau Isoscape"
author: "CB Wall"
date: "9/11/2019"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 4
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '4'
editor_options: 
  chunk_output_type: inline
---

```{r setup, message=FALSE, warning=FALSE, echo=FALSE}
require("knitr")
knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.align='center')

# load packages
library("RgoogleMaps")
library("SDMTools")
library("rgdal")
library("ggmap")
library("gridExtra")
```

## Background  
**Hakalau National Forest Wildlife Refuge Isoscapes**  
Data collected 20-21 August 2019 in Hakalau Forest, Hawai'i Island. Soil and plant samples collected from remnant ko forests (*RK sites*) and those afforested/planted after a century of deforestation from cattle grazing (*AK sites*). Target engineers of forest habitats are *Acacia koa* (common name: koa), an indigenous hardwood species that forms native forest canopies and associates with nitrogen fixing bacteria. Understory species include *Rubus argutus* -- an invasive species (common name: sawtooth blackberry) -- and an endemic species *Rubus hawaiiensis* (common name: 'Ākala); both are members of the Roseaceae family.  

Soil samples were collected in a spatially explicit fashion (every 4 or 5m) with a 20 x 35m superplot, consisting of thirty-five 4 x 5m subplots. This data was used to create a C- and N-stable isotope landscape, hereafter 'isoscape'. Koa and *Rubus* spp. were collected within each 4 x 5m sublot.


### Site Maps
```{r, site maps}

# load data
Hak.gps<-read.csv("data/Hakalau.gps.csv")
API<-read.csv("data/API_key.csv")
API.key<-API[1,1]

ggplot(Hak.gps, aes(x = longitude, y = latitude)) +
  coord_quickmap() +
  geom_point()

######## using ggmap
register_google(key=API.key)

Hak.rev=c(-155.317, 19.82158002)
map2<-get_map(Hak.rev, 
                      zoom=17, 
                      scale = 2, 
                      maptype= "satellite",
                      source="google", extent= "device", legend="topright")

png(file= "figures/sitemap.test.png", height=5, width=5, units="in", res=300)
par(oma=c(3,3,0,0))
ggmap(map2)+
  geom_point(aes(x=longitude, y=latitude), data=Hak.gps, alpha=0.5, color="dodgerblue", size=1)+
  labs(x="Longitude", y="Latitude")
dev.off()

# AK or RK only
AK<- Hak.gps[(Hak.gps$Site=="AK"),]
RK<- Hak.gps[(Hak.gps$Site=="RK"),]

####### AK map
AK.gps<-c(-155.3189, lat=19.8219)
mapAK<-get_map(AK.gps,
              zoom=20, 
              scale = 2, 
              maptype= "satellite",
              source="google", extent= "device", legend="topright")
AK.map<-ggmap(mapAK, group=type)+
  geom_point(aes(x=longitude, y=latitude, shape=type, color=type), data=AK, alpha=0.5, size=3)+
  theme(legend.position="top",
    legend.text = element_text(color = "white"),
    legend.title = element_text(color = "white"),
    legend.key = element_rect(fill = "white")) +
  guides(colour= guide_legend(override.aes = list(color = "white"))) +
  scale_color_manual(values=c('red','mediumseagreen'))+
  labs(x="Longitude", y="Latitude") + ggtitle("AK-Aforested")


####### RK map
RK.gps<-c(-155.315, 19.8216)
mapRK<-get_map(RK.gps,
               zoom=20, 
               scale = 2, 
               maptype= "satellite",
               source="google", extent= "device", legend="topright")
RK.map<-ggmap(mapRK, group=type)+
  geom_point(aes(x=longitude, y=latitude, shape=type, color=type), data=RK, alpha=0.5, size=3)+
  scale_color_manual(values=c('red','mediumseagreen', "dodgerblue"))+
  scale_shape_manual(values=c(16, 17, 3))+
  theme(legend.position="top",
        legend.title=element_blank(), 
        legend.key = element_rect(fill = "white"))+
  labs(x="Longitude", y="") + ggtitle("RK-Remnant")

### export it
png(file= "figures/RK.AK.sitemaps.png", height=5, width=9, units="in", res=300)
grid.arrange(AK.map, RK.map, ncol = 2)
dev.off()

```



```{r}


```

```{r}


```


```{r}

```

