---
title: "Hakalau Isoscape"
author: "CB Wall"
date: "9/11/2019"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 4
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '4'
editor_options: 
  chunk_output_type: console
---

```{r global options, results="hide", warning=FALSE, message=FALSE}
if (!require('knitr')) install.packages('knitr'); library('knitr')
knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.align='center')

# Load in packages
if (!require("pacman")) install.packages("pacman"); library(pacman) # for rapid install if not in library

# load packages
pacman::p_load("RgoogleMaps", "SDMTools", "rgdal", "ggmap", "gridExtra", "automap", "cowplot","sp", "gstat", "plotrix", "dplyr", "ggplot2", "scales","magrittr","plyr","effects", "ggpubr", "sjstats", "RColorBrewer")

```

## Background  
**Hakalau National Forest Wildlife Refuge Isoscapes**  
Data collected 20-21 August 2019 in Hakalau Forest, Hawai'i Island. Soil and plant samples collected from  afforested/restored plots after a century of deforestation from cattle grazing (*AK sites*) and remnant mixed koa-ohia forests (*RK sites*). Target engineers of forest habitats are *Acacia koa* (Hawaiian name: *Koa*), an indigenous hardwood species that forms native forest canopies and associates with nitrogen fixing bacteria. Understory species include *Rubus argutus* -- an invasive species (common name: sawtooth blackberry) -- and an endemic species *Rubus hawaiiensis* (common name: 'Ākala); both are members of the Roseaceae family. Noteably, *R. argutus* was found more in the remnant plot (i.e., RK) and the indigenous *R. hawaiiensis* was common in the restored plot (i.e, AK). 

Soil samples were collected in a spatially explicit fashion (every 4 or 5m) with a 20 x 35m superplot, consisting of thirty-five 4 x 5m subplots. This data was used to create a δ^13^C and δ^15^N stable isotope landscape, hereafter 'isoscape'. Koa and *Rubus* spp. were collected within each 4 x 5m sublot.  

*Plot densities were...*  
AK: Koa (18 trees) Rubus (14 plants) per 20 x 35m plot= *0.026 Koa m^-2^, 0.020 Rubus m^-2^*
RK: Koa (10 trees) Rubus (28 plants) per 20 x 35m plot= *0.014 Koa m^-2^, 0.040 Rubus m^-2^*

### Site Maps
```{r, site maps, fig.cap="Figure 1. Site map of (a) the Hawaiian Island archipelago, (b) sampling plot in Hakalau Forest Refuge, and (c) plot layouts, highlighting *Acacia koa* density.", fig.align='center'}

# load data
Hak.gps<-read.csv("data/Hakalau.gps.csv")
API<-read.csv("data/API_key.csv")
API.key<-API[1,1]

ggplot(Hak.gps, aes(x = longitude, y = latitude)) +
  coord_quickmap() +
  geom_point()

######## using ggmap
register_google(key=API.key)

########
#Hawaiian islands Map
########
hi_map<-get_map(location=c(-157,20.5),zoom=7,maptype="satellite",color="color")
hi_map_for_man <- ggmap(hi_map) +
  geom_point(aes(x = -155.320, y = 19.83), pch=23,colour="black",fill="red", size = 2, stroke=0.5) +
  xlab("") + ylab("Latitude") +
  scale_y_continuous(limits=c(18.8, 22.2))+
  scale_x_continuous(limits=c(-160.2, -154)) +
  theme(axis.text=element_text(colour="black",size=5),
        axis.title=element_text(colour="black",size=8),
        plot.margin = unit(c(0, 0.5, 0, 0.7), "cm")) +
  ggsn::scalebar(x.min=-160.2, x.max=-154, y.min=18.8, y.max=22.2, dist=50, dist_unit="km", transform=TRUE,
                 st.bottom=FALSE, st.size=2, box.fill=c("gray50", "white"), model="WGS84",st.color="white", border.size=0.5)


##########
# site map plot showing distance from plots
##########
Hak.rev=c(-155.317, 19.82158002)

map2<-get_map(Hak.rev, 
                      zoom=16, 
                      scale = 2, 
                      maptype= "satellite",
                      source="google", extent= "device", legend="topright")

## site single point only
site.only<-Hak.gps[c(1,5),]
Hakalau.site<-
  ggmap(map2)+
  geom_point(aes(x=longitude, y=latitude), data=site.only, alpha=0.5, color="white", size=4)+
  labs(x="", y="") +
  scale_y_continuous(limits=c(19.81785, 19.8242))+
  scale_x_continuous(limits=c(-155.3223, -155.311)) +
  theme(text = element_text(size=6),
       plot.margin = unit(c(0.2, 0.5, 0.2, 0.2), "cm")) +
  annotate("text", x=-155.3187, y=19.8225, label= "AK", size=3, col="white") +
  annotate("text", x=-155.31478, y=19.8224, label= "RK", size=3, col="white") +
ggsn::scalebar(x.min=-155.314, x.max=-155.311, y.min=19.818, y.max=19.824, dist=100, 
               dist_unit="m", transform=TRUE,
              st.bottom=FALSE, st.size=2, box.fill=c("gray50", "white"), model="WGS84",st.color="white", border.size=0.5)
Hakalau.site


##########
# Site plots: AK or RK only
##########
AK<- Hak.gps[(Hak.gps$Site=="AK"),]
RK<- Hak.gps[(Hak.gps$Site=="RK"),]

####### AK map
AK.gps<-c(-155.3189, lat=19.8219)
mapAK<-get_map(AK.gps,
              zoom=19, 
              scale = 2, 
              maptype= "satellite",
              source="google", extent= "device", legend="topright")

AK.map<-ggmap(mapAK, group=type)+
  geom_point(aes(x=longitude, y=latitude, shape=type, color=type), data=AK, alpha=0.5, size=3)+
  scale_y_continuous(limits=c(19.82165, 19.82225))+
  scale_x_continuous(limits=c(-155.3194, -155.3183)) +
  theme(legend.position="top",
    legend.text = element_text(color = "white"),
    legend.title = element_text(color = "white"),
    legend.key = element_rect(fill = "white"),
    plot.margin = unit(c(0, 0.5, 0, 0), "cm"),
    text = element_text(size=8)) +
  guides(colour= guide_legend(override.aes = list(color = "white"))) +
  scale_color_manual(values=c('red','mediumseagreen'))+
  labs(x="Longitude", y="Latitude") +
  annotate("text", x=-155.31925, y=19.82225, label= "Restored Forest (AK)", size=3, col="white")+
  ggsn::scalebar(x.min=-155.3183, x.max=-155.3187, y.min=19.82165, y.max=19.82225, dist=10, 
               dist_unit="m", transform=TRUE,
              st.bottom=FALSE, st.size=2, box.fill=c("gray50", "white"), model="WGS84", st.color="white", border.size=0.5)


####### RK map
RK.edit<-RK[-12,] # remove "proximate" Koa
RK.edit[9,3]<-"koa" # rename to be "koa"

RK.gps<-c(-155.315, 19.8216)
mapRK<-get_map(RK.gps,
               zoom=19, 
               scale = 2, 
               maptype= "satellite",
               source="google", extent= "device", legend="topright")

RK.map<-ggmap(mapRK, group=type)+
  geom_point(aes(x=longitude, y=latitude, shape=type, color=type), data=RK.edit, alpha=0.5, size=3)+
  scale_color_manual(values=c('red','mediumseagreen', "dodgerblue"))+
  scale_shape_manual(values=c(16, 17, 3)) +
  scale_y_continuous(limits=c(19.82120, 19.82195))+
  scale_x_continuous(limits=c(-155.31575, -155.3144)) +
  theme(legend.position="top",
        legend.title=element_blank(), 
        legend.key = element_rect(fill = "white"),
        plot.margin = unit(c(0, 0.5, 0, 0), "cm"), 
        text = element_text(size=8)) +
  labs(x="Longitude", y="") +
  annotate("text", x=-155.3156, y=19.82193, label= "Remnant Forest (RK)", size=3, col="white")+
  ggsn::scalebar(x.min=-155.3148, x.max=-155.3144, y.min=19.8212, y.max=19.8219, dist=10, 
               dist_unit="m", transform=TRUE,
              st.bottom=FALSE, st.size=2, box.fill=c("gray50", "white"), model="WGS84", st.color="white", border.size=0.5)


site.plots<-plot_grid(hi_map_for_man, Hakalau.site, AK.map, RK.map, 
          labels=c('a', 'b', 'c', 'd'), label_size=8, hjust=-1, vjust= 2, ncol=2, nrow=2)

### export it
pdf(file= "figures/Fig1.sitemaps.pdf", height=7, width=8)
site.plots
dev.off()

```

#### Ecology and Isotopes data  
- DBH calculations  
- data mutating  
- test scatters of all samples by habitat  
```{r, Hak data, fig.show=FALSE}
Hak.df<-read.csv("data/Hakalau.isotopes.csv")

########## DBH 
DBH.df<- Hak.df %>%
  select(c(DBH..cm.1, DBH..cm.2, DBH..cm.3, DBH..cm.4, DBH..cm.5))

# replace NAs with zeros
DBH.df<- DBH.df %>%
  mutate_each(funs(replace(., which(is.na(.)), 0)))

# if only one trunk, then make new column and report this value
DBH.df <- DBH.df %>%
  mutate(DBH.1trunk = ifelse(DBH..cm.2>"0", 0, DBH..cm.1))

# if multiple trunks, sum of squares and square root all trunks
# and plug in the DBH for single trunks
DBH.df <- DBH.df %>%
  mutate(DBH.Total = ifelse(DBH.1trunk=="0",  
                     sqrt(DBH..cm.1^2 + DBH..cm.2^2 + DBH..cm.3^2 + DBH..cm.4^2 + DBH..cm.5^2), DBH.1trunk))

# replace zeros with NA now that total DBH has been calculated
DBH.df$DBH.Total<-as.numeric(ifelse(DBH.df$DBH.Total=="0", "NA", DBH.df$DBH.Total))

############ Canopy diameter
# use area of an elipse, a x b x π, where a and b are major and minor radius
Can.df<-Hak.df %>%
  select(c(canopy.0..m, canopy.90..m, canopy.180..m, canopy.270..m))

# Major radius is the average of 0 and 180 degrees
Can.df<- Can.df %>% 
  mutate(Maj.rad = rowMeans(cbind(canopy.0..m, canopy.180..m), na.rm=T))

# Minor radius is the average of 90 and 270 degrees
Can.df<- Can.df %>% 
  mutate(Min.rad = rowMeans(cbind(canopy.90..m, canopy.270..m), na.rm=T))

# Elipse area
Can.df<- Can.df %>% 
  mutate(Canopy.area = ifelse(Maj.rad>0, Maj.rad*Min.rad*3.14, "NA"))

########################
# combine DBH.Total and Canopy.area with isotope dataframe and reduce columns
Hak.df$DBH.Total<-DBH.df$DBH.Total
Hak.df$Canopy.area<-Can.df$Canopy.area

# Hak.data is the full dataframe
Hak.data<- Hak.df %>%
  select(c(Plot, Position.point, Sample, DBH.Total, Canopy.area, d15N, d13C, N..percent, Total.N..mmol.gdw, C..percent, Total.C..mmol.gdw, C.N))

########################
########################
########################

# test plots

### Scatter of d13C vs. d15N
iso.scatter<-ggplot(data=Hak.data, aes(x=d13C, y=d15N, color=Plot, shape=Sample))+
  geom_point(cex=1.5) +
  xlab(expression(paste(delta^{13}, C, " (\u2030, V-PDB)"))) +
  ylab(expression(paste(delta^{15}, N, " (\u2030, air)"))) 

### Scatter of total C vs. total N
total.nut<-ggplot(data=Hak.data, aes(x=Total.C..mmol.gdw, y=Total.N..mmol.gdw, color=Plot, shape=Sample))+
  geom_point(cex=1.5) +
  xlab("total Carbon") +
  ylab("total Nitrogen")

#  figures together
plot_grid(iso.scatter, total.nut,
     labels=c('a', 'b'), label_size=10, hjust=-1, vjust= 3, ncol=2, nrow=1)
```
  
#### Rubus and Koa relationship  
- plotting the relationship between Rubus and Koa d15N  
- took average of 3 closest Koa trees to each Rubus plant  
- plot Rubus d15N against Koa d15N (mean of 3 trees)  

```{r rubus and koa, fig.cap="Figure 2a. Relationship between d15N-Koa and 15N-Rubus at two sampling sites.", fig.align='center'}
Ndfa<-read.csv("data/Ndfa.Rub.csv")

# rename the d15N so we know it is for rubus
Ndfa$d15N.Rub<-Ndfa$d15N
Ndfa$d15N.Koa<-Ndfa$mean.d15N.Ndfa

####################
#################### relationship between koa and rubus d15N?
# Yes, but a bit stronger in the RK site
mod.AK<-lm(Ndfa[(Ndfa$Plot=="AK"),]$d15N.Rub~Ndfa[(Ndfa$Plot=="AK"),]$d15N.Koa); anova(mod.AK)
int.rubAK<-coef(summary(mod.AK))[1] #intercept
slop.rubAK<-coef(summary(mod.AK))[2] #slope
                                  
mod.RK<-lm(Ndfa[(Ndfa$Plot=="RK"),]$d15N.Rub~Ndfa[(Ndfa$Plot=="RK"),]$d15N.Koa); anova(mod.RK)
int.RKoa<-coef(summary(mod.RK))[1] #intercept
slop.RKoa<-coef(summary(mod.RK))[2] #slope

#### plot
BW.back<-theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black", size=0.2))

Rub.Koa.d15N.plot<-ggplot(data=Ndfa, aes(x=d15N.Koa, y=d15N.Rub, color=Plot))+
  geom_point(aes(color=Plot), alpha=.8) + 
  scale_color_manual(values=c('gray30', "gray70")) +
  coord_equal() + theme_bw() +
  xlab(expression(paste(italic("Acacia koa "), delta^{15}, N, " (\u2030)"), sep=""))+
  ylab(expression(paste(italic("Rubus"), " spp. ", delta^{15}, N, " (\u2030)"), sep="")) +
  guides(colour=guide_legend(override.aes = list(stroke=1, fill=NA, alpha=1))) +
  geom_smooth(method = "lm", alpha = .15, size=0.6) +
  theme(legend.title = element_blank(),
        axis.text=element_text(size=6),
        axis.title=element_text(size=10))+
  annotate(geom="text", x=2.7, y=0.8, label=expression(paste("AK ", italic("p"), "=0.056")), size=2) +
  annotate(geom="text", x=2.7, y=0.6, label=expression(paste("RK ", italic("p="), bold("0.006"))), size=2) +
  BW.back

Rub.Koa.d15N.plot
dev.copy(pdf, "figures/Fig 3.Koa.Rub.regr.pdf", width=4, height=4, encod="MacRoman")
dev.off()
```
  

#### DBH, Canopy  d15N models  
- plot against d15N values with model curves
- scatters with models
- d15N differ by PLOT (AK vs RK), no other effects or relationship with DBH or Canopy

```{r, d15N and DBH Canopy, eval=FALSE}
# ** No relationship between d15N and DBH or Canopy at plot level**
# ** Effects at PLOTs on d15N (lower in AK) **


######## plot of d15N and DBH, and d15N and Canopy area

# all samples
par(mfrow=c(1,2), mar=c(4,5,1,1))

#### DBH
mod.DBH<-lm(d15N~DBH.Total*Plot, data=Hak.data) # no interaction effect

mod.DBH<-lm(d15N~DBH.Total+Plot, data=Hak.data)
int<-coef(summary(mod.DBH))[1]
DBH.slope<-coef(summary(mod.DBH))["DBH.Total",1] # DBH (slope for all figures)
RK.int<-coef(summary(mod.DBH))["PlotRK",1] # RK, AK = 0


# no significant relationship to d15N but a trend for DECREASE with DBH
plot(d15N~DBH.Total, data=Hak.data,
    col=c("coral","dodgerblue")[as.factor(Plot)],
    pch=c(1,2)[as.factor(Plot)],
    ylab=expression(paste(delta^{15}, N, " (\u2030, air)")),
    xlab="DBH (cm)",
    ylim=c(0,5), xlim=c(0,60))
    ablineclip(int, DBH.slope, col="coral", lwd=2, 
               x1 = min(Hak.data$DBH.Total[Hak.data$Plot=="AK"], na.rm=T), 
               x2 = max(Hak.data$DBH.Total[Hak.data$Plot=="AK"], na.rm=T)) # AK model
    ablineclip((int+ RK.int), DBH.slope, col="dodgerblue", lwd=2, 
               x1 = min(Hak.data$DBH.Total[Hak.data$Plot=="RK"], na.rm=T), 
               x2 = max(Hak.data$DBH.Total[Hak.data$Plot=="RK"], na.rm=T)) # RK model

    
#### Canopy
mod.Can<-lm(d15N~Canopy.area*Plot, data=Hak.data) # no interaction

mod.Can<-lm(d15N~Canopy.area+Plot, data=Hak.data)
int<-coef(summary(mod.Can))[1]
Can.slope<-coef(summary(mod.Can))["Canopy.area",1] # DBH (slope for all figures)
RK.int<-coef(summary(mod.Can))["PlotRK",1] # RK, AK = 0

# no significant relationship to d15N but a trend for INCREASE with canopy area
plot(d15N~Canopy.area, data=Hak.data, yaxt="n",
     col=c("coral","dodgerblue")[as.factor(Plot)],
     pch=c(1,2)[as.factor(Plot)],
     xlab=expression(paste("Canopy (m"^2,")")),
     ylab="", 
     ylim=c(0,5), xlim=c(0,80))
     axis(side=2, labels=F)
     ablineclip(int, Can.slope, col="coral", lwd=2, 
               x1 = min(Hak.data$Canopy.area[Hak.data$Plot=="AK"], na.rm=T), 
               x2 = max(Hak.data$Canopy.area[Hak.data$Plot=="AK"], na.rm=T)) # AK model
     ablineclip((int+ RK.int), Can.slope, col="dodgerblue", lwd=2, 
               x1 = min(Hak.data$Canopy.area[Hak.data$Plot=="RK"], na.rm=T), 
               x2 = max(Hak.data$Canopy.area[Hak.data$Plot=="RK"], na.rm=T)) # RK model
     legend("topright", legend=levels(Hak.data$Plot), box.lty=0, bg="transparent", lty=1, pch=c(1,2), 
            col=c("coral", "dodgerblue"), cex=1)

dev.copy(pdf, "figures/d15N.DBH.Canopy.pdf", width=7, height=5, encod="MacRoman")
dev.off()
```
  
With Plot being most important effect, divide DBH and Canopy to be plot mean +/- SE  
*Supplemental Figure 1. with afforested and remnant forests*   
- DBH and Canopy figure as bar chart and means
- using Mann-Whitney nonparametric tests: significant plot effects
- greater DBH (AK), trend for greater Canopy (RK)  
  
```{r DBH and Canopy, fig.show='hide'}
## Models
mod.DBH<-lm(DBH.Total~Plot, data=Hak.data); anova(mod.DBH) # signif
mod.Can<-lm(Canopy.area~Plot, data=Hak.data); anova(mod.Can) # not signif, a bit wobbly in assumptions

# diagnostic plots
for(i in c(4:5)){
  Y<-Hak.data[,i]
  full<-lm(Y~Plot, data=Hak.data, na.action=na.exclude)
  R <- resid(full) #save glm residuals
  op<-par(mfrow = c(2,3), mar=c(5,4,1,2), pty="sq")
  plot(full, add.smooth = FALSE, which=1)
  QQ <- qqnorm(R, main = colnames(Hak.data)[i]) 
  QQline <- qqline(R)
  hist(R, xlab="Residuals", main = colnames(Hak.data)[i])
  plot(Hak.data$Plot, R, xlab=colnames(Hak.data)[i], ylab="Residuals")
}

# Mann-Whitney U-Test
mwu(Hak.data, DBH.Total, Plot) # signif
mwu(Hak.data, Canopy.area, Plot) # not
```

```{r mean plots for DBH and Canopy, Fig.cap="Supplementary Figure 1. Mean +/- SE diameter at breast height (DBH) and *A. koa* canopy area for two afforested and remnant forest plots.", fig.align='center'}
#### plots
DBH.m<-aggregate(DBH.Total~Plot, data=Hak.data, mean)
Can.m<-aggregate(Canopy.area~Plot, data=Hak.data, mean)
canopy.mean<-aggregate(Canopy.area~Plot, data=Hak.data, sum) 

DBH.SE<-aggregate(DBH.Total~Plot, data=Hak.data, std.error)
Can.SE<-aggregate(Canopy.area~Plot, data=Hak.data, std.error)

DBH.n<-aggregate(DBH.Total~Plot, data=Hak.data, length)
Can.n<-aggregate(Canopy.area~Plot, data=Hak.data, length)

mean.eco<-cbind(DBH.m, Can.m[2], DBH.SE[2], Can.SE[2]); colnames(mean.eco)<-c("Plot", "DBH", "Can", "D.SE", "C.SE")

pd <- position_dodge(0.7) #offset for error bars

BW.back2<-theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black", size=0.4))

### DBH plot of mean/SE by plot
DBH.plot<-ggplot(data=mean.eco, aes(x=Plot, y=DBH))+
  geom_bar(stat="identity", position = pd, width=0.7, fill="gray70")+
  geom_errorbar(aes(ymin=DBH-D.SE, ymax=DBH+D.SE), size=.5, width=0, position=pd) +
  xlab("Plots") +
  ylab("dbh (cm)") + 
  theme(legend.title = element_blank(),
        axis.text=element_text(size=6),
        axis.title=element_text(size=8))+
  annotate(geom="text", x=1.5, y=35, label=expression(paste(bold("*"))), size=4) +
  BW.back2

### Canopy plot of mean/SE by plot
Canopy.plot<-ggplot(data=mean.eco, aes(x=Plot, y=Can))+
  geom_bar(stat="identity", position = pd, width=0.7, fill="gray70")+
  geom_errorbar(aes(ymin=Can-C.SE, ymax=Can+C.SE), size=.5, width=0, position=pd) +
  xlab("Plots") +
  ylab(expression(paste("Canopy (m"^2,")"))) + 
  theme(legend.title = element_blank(),
        axis.text=element_text(size=6),
        axis.title=element_text(size=8))+
  BW.back2

#  figures together
plot_grid(DBH.plot, Canopy.plot,
     labels=c('a', 'b'), label_size=8, hjust=-1, vjust= 3, ncol=2, nrow=1)
dev.copy(pdf, "figures/Supp.Fig1.DBHCanopy.pdf", width=4, height=3.5, encod="MacRoman")
dev.off()

```
  
Run statistics on the responses (dd15N, 13C, tissue C and N)
```{r, stats on responses, results='hide', fig.show=FALSE}

# models

## tests for assumptions
for(i in c(6:12)){
  Y<-Hak.data[,i]
  full<-lm(Y~Plot*Sample, data=Hak.data, na.action=na.exclude)
  R <- resid(full) #save glm residuals
 
  op<-par(mfrow = c(2,3), mar=c(5,4,1,2), pty="sq")
  plot(full, add.smooth = FALSE, which=1)
  QQ <- qqnorm(R, main = colnames(Hak.data)[i]) 
  QQline <- qqline(R)
  hist(R, xlab="Residuals", main = colnames(Hak.data)[i])
  plot(Hak.data$Plot, R, xlab=colnames(Hak.data)[i], ylab="Residuals")
  plot(Hak.data$Sample, R, xlab=colnames(Hak.data)[i], ylab="Residuals")
}

########## Separate dataframes
soil<-Hak.data[(Hak.data$Sample=="Soil"),]
Koa<-Hak.data[(Hak.data$Sample=="Koa"),]
Rubus<-Hak.data[(Hak.data$Sample=="RUBARG" | Hak.data$Sample=="RUBHAW"),]
AK.df<-Hak.data[(Hak.data$Plot=="AK"),]; AK.df.sansoil<-AK.df[!(AK.df$Sample=="Soil"),]
RK.df<-Hak.data[(Hak.data$Plot=="RK"),]; RK.df.sansoil<-RK.df[!(RK.df$Sample=="Soil"),]

```
  
**d15N model**
```{r, d15N}
######## Nitrogen isotopes
d15N.soil<-lm(d15N~Plot, data=soil); anova(d15N.soil) # soil signif.
plot(allEffects(d15N.soil))

d15N.Koa<-lm(d15N~Plot, data=Koa); anova(d15N.Koa) # Koa signif.
plot(allEffects(d15N.Koa))

d15N.Rubus<-lm(d15N~Plot, data=Rubus); anova(d15N.Rubus) # Rubus signif.
plot(allEffects(d15N.Rubus))

d15N.AK<-lm(d15N~Sample, data=AK.df.sansoil); anova(d15N.AK) #only koa vs. rubus comparision, signif.
plot(allEffects(d15N.AK)) 

d15N.RK<-lm(d15N~Sample, data=RK.df.sansoil); anova(d15N.RK) #only koa vs. rubus comparision, signif.
plot(allEffects(d15N.RK))

```
  
**d13C model**
```{r, d13C}
######## ######## 
######## Carbon isotopes
d13C.soil<-lm(d13C~Plot, data=soil); anova(d13C.soil) # soil signif.
plot(allEffects(d13C.soil))

d13C.Koa<-lm(d13C~Plot, data=Koa); anova(d13C.Koa) # Koa signif.
plot(allEffects(d13C.Koa))

d13C.Rubus<-lm(d13C~Plot, data=Rubus); anova(d13C.Rubus) # Rubus signif.
plot(allEffects(d13C.Rubus))

d13C.AK<-lm(d13C~Sample, data=AK.df.sansoil); anova(d13C.AK) #only koa vs. rubus comparision
plot(allEffects(d13C.AK))

d13C.RK<-lm(d13C~Sample, data=RK.df.sansoil); anova(d13C.RK) #only koa vs. rubus comparision
plot(allEffects(d13C.RK))


```
 
**Nitrogen content model**
```{r, N concent}
####### Nitrogen content
TN.soil<-lm(Total.N..mmol.gdw~Plot, data=soil); anova(TN.soil) # soil signif.
plot(allEffects(TN.soil))

TN.Koa<-lm(Total.N..mmol.gdw~Plot, data=Koa); anova(TN.Koa) # Koa signif.
plot(allEffects(TN.Koa))

TN.Rubus<-lm(Total.N..mmol.gdw~Plot, data=Rubus); anova(TN.Rubus) # Rubus signif.
plot(allEffects(TN.Rubus))

N.AK<-lm(Total.N..mmol.gdw~Sample, data=AK.df.sansoil); anova(N.AK) #only koa vs. rubus comparision, signif.
plot(allEffects(N.AK)) 

N.RK<-lm(Total.N..mmol.gdw~Sample, data=RK.df.sansoil); anova(N.RK) #only koa vs. rubus comparision, signif.
plot(allEffects(N.RK))

```
 
**Carbon content model**
```{r, C concent}
######## ######## 
######## Carbon content
TC.soil<-lm(Total.C..mmol.gdw~Plot, data=soil); anova(TC.soil) # soil signif.
plot(allEffects(TC.soil))

TC.Koa<-lm(Total.C..mmol.gdw~Plot, data=Koa); anova(TC.Koa) # Koa signif.
plot(allEffects(TC.Koa))

TC.Rubus<-lm(Total.C..mmol.gdw~Plot, data=Rubus); anova(TC.Rubus) # Rubus signif.
plot(allEffects(TC.Rubus))

C.AK<-lm(Total.C..mmol.gdw~Sample, data=AK.df.sansoil); anova(C.AK) #only koa vs. rubus comparision, signif.
plot(allEffects(C.AK)) 

C.RK<-lm(Total.C..mmol.gdw~Sample, data=RK.df.sansoil); anova(C.RK) #only koa vs. rubus comparision, signif.
plot(allEffects(C.RK))

```
  
**C:N**
```{r, CN values}
####### C:Ncontent
CN.soil<-lm(C.N~Plot, data=soil); anova(CN.soil) # not signif.

CN.Koa<-lm(C.N~Plot, data=Koa); anova(CN.Koa) # not signif.

CN.Rubus<-lm(C.N~Plot, data=Rubus); anova(CN.Rubus) # not signif

CN.AK<-lm(C.N~Sample, data=AK.df.sansoil); anova(CN.AK) #only koa vs. rubus comparision, signif.
plot(allEffects(CN.AK)) 

CN.RK<-lm(C.N~Sample, data=RK.df.sansoil); anova(CN.RK) #only koa vs. rubus comparision, not signif
plot(allEffects(CN.RK))

```

*Figure 2*   
data means and SE for plots  
- bar plots of d13C, d15N, TC and TN for all 4 samples types with asterisks for significant effects

```{r, plots of isotope data, Fig.cap="Supplementary Figure 2. Mean +/- SE of sample (a) d13C, (b) d15N, (c) total carbon, and (d) total nitrogen.", fig.align='center'}
## make some means and SE
d13C.m<-aggregate(d13C~Plot+Sample, data=Hak.data, mean)
d15N.m<-aggregate(d15N~Plot+Sample, data=Hak.data, mean)
TN.m<-aggregate(Total.N..mmol.gdw~Plot+Sample, data=Hak.data, mean)
TC.m<-aggregate(Total.C..mmol.gdw~Plot+Sample, data=Hak.data, mean)
CN.m<-aggregate(C.N~Plot+Sample, data=Hak.data, mean)

d13C.SE<-aggregate(d13C~Plot+Sample, data=Hak.data, std.error)
d15N.SE<-aggregate(d15N~Plot+Sample, data=Hak.data, std.error)
TN.SE<-aggregate(Total.N..mmol.gdw~Plot+Sample, data=Hak.data, std.error)
TC.SE<-aggregate(Total.C..mmol.gdw~Plot+Sample, data=Hak.data, std.error)
CN.SE<-aggregate(C.N~Plot+Sample, data=Hak.data, std.error)

d13C.n<-aggregate(d13C~Plot+Sample, data=Hak.data, length)
d15N.n<-aggregate(d15N~Plot+Sample, data=Hak.data, length)

mean.data<- cbind(d13C.m, d15N.m[3], TN.m[3], TC.m[3], d13C.SE[3], d15N.SE[3], TN.SE[3], TC.SE[3], CN.m[3], CN.SE[3])
colnames(mean.data)<-c("Plot", "Sample", "d13C.mean", "d15N.mean", "TN.mean", "TC.mean", "d15N.SE", "d13C.SE", "TN.SE", "TC.SE", "CN.mean", "CN.SE")

# rename Rubus species for plotting
levels(mean.data$Sample)[levels(mean.data$Sample)=="RUBARG"] <- "Rubus spp"
levels(mean.data$Sample)[levels(mean.data$Sample)=="RUBHAW"] <- "Rubus spp"

mean.data$Sample<-factor(mean.data$Sample, levels=c("Soil", "Koa", "Rubus spp"))

### make mean plots
pd <- position_dodge(0.7) #offset for error bars

### d13C
d13C.plot<-ggplot(data=mean.data, aes(x=Sample, y=d13C.mean, fill=Plot, group=Plot))+
  geom_bar(stat="identity", position = pd, width=0.7)+
  geom_errorbar(aes(ymin=d13C.mean-d13C.SE, ymax=d13C.mean+d13C.SE), size=.3, width=0, position=pd) +
  scale_fill_manual(values=c('gray65', "gray89")) +
  xlab("Sample Type") + 
  annotate(geom="text", x=1, y=-27.5, label=expression(paste(bold("*"))), size=4) +
  annotate(geom="text", x=2, y=-33, label=expression(paste(bold("*"))), size=4) +
  annotate(geom="text", x=3, y=-33, label=expression(paste(bold("*"))), size=4) +
  theme(text = element_text(size=8)) +
  ylab(expression(paste(delta^{13}, C, " (\u2030, V-PDB)"))) + BW.back2

### d15N
d15N.plot<-ggplot(data=mean.data, aes(x=Sample, y=d15N.mean, fill=Plot, group=Plot))+
  geom_bar(stat="identity", position = pd, width=0.7)+ ylim(c(0,8)) +
  geom_errorbar(aes(ymin=d15N.mean-d15N.SE, ymax=d15N.mean+d15N.SE), size=.3, width=0, position=pd) +
  scale_fill_manual(values=c('gray65', "gray89")) +
  xlab("Sample Type") + 
  theme(text = element_text(size=8)) +
  annotate(geom="text", x=2, y=2.8, label=expression(paste(bold("*"))), size=4) +
  annotate(geom="text", x=3, y=3.3, label=expression(paste(bold("*"))), size=4) +
  ylab(expression(paste(delta^{15}, N, " (\u2030, air)"))) + BW.back2

### total Carbon
TC.plot<-ggplot(data=mean.data, aes(x=Sample, y=TC.mean, fill=Plot, group=Plot))+
  geom_bar(stat="identity", position = pd, width=0.7)+ ylim(c(0, 60)) +
  geom_errorbar(aes(ymin=TC.mean-TC.SE, ymax=TC.mean+TC.SE), size=.3, width=0, position=pd) +
  scale_fill_manual(values=c('gray65', "gray89")) +
  xlab("Sample Type") + 
  theme(text = element_text(size=8)) +
  annotate(geom="text", x=1, y=25, label=expression(paste(bold("*"))), size=4) +
  annotate(geom="text", x=3, y=42, label=expression(paste(bold("*"))), size=4) +
  ylab("Total Carbon (mmol/gdw)") + BW.back2

### total Nitrogen
TN.plot<-ggplot(data=mean.data, aes(x=Sample, y=TN.mean, fill=Plot, group=Plot))+
  geom_bar(stat="identity", position = pd, width=0.7)+
  geom_errorbar(aes(ymin=TN.mean-TN.SE, ymax=TN.mean+TN.SE), size=.3, width=0, position=pd) +
  scale_fill_manual(values=c('gray65', "gray89")) +
  xlab("Sample Type") + 
  theme(text = element_text(size=8)) +
  ylab("Total Nitrogen (mmol/gdw)") + BW.back2

# get the legend, # create some space to the left of the legend
bar.legend <- get_legend(
  TN.plot + theme(legend.box.margin = margin(0, 0, 0, 12)) + 
    theme(legend.key.size = unit(0.3, "cm")))

#  figures together
bar.plots<-plot_grid(d15N.plot + theme(legend.position = "none"), 
                     d13C.plot + theme(legend.position = "none"),
                     TN.plot + theme(legend.position = "none"),
                     TC.plot + theme(legend.position = "none"), 
                     nrow=1, ncol=4, labels=c('a', 'b', 'c', 'd'), 
          label_size=8, hjust=-1, vjust= 3)

plot_grid(bar.plots, bar.legend, rel_widths = c(8, 1)) # legend  1/8 size as first obj.
dev.copy(pdf, "figures/Fig 2.mean.data.sample.pdf", width=7, height=3, encod="MacRoman")
dev.off()
```
  
###ISO-KRIG
Give the data a "krig-over".  
  
First, make krigs of the full plot and all data layers
```{r, krig setup}
# load data
krig<-read.csv("data/krig.matrix.csv") # matrix of points for isoscape grid

# merge kriging data with isotope data
data.merge<-as.data.frame(join_all(list(Hak.data, krig), by = "Position.point", type='full'))

data.merge<-data.merge[c(-167:-194),] # drop NAs where no samples taken
scape.data<-data.merge

write.csv(scape.data, "output/scape.data.csv")

#################
```

####d15N isoscape   
setup
```{r, krig d15N}
################# d15N krigs
################# make site maps based on bubble plots

# just AK site
krig.AK<- scape.data %>% filter(Plot=="AK")
krig.AK$Sample<-droplevels(krig.AK$Sample)

# modify one point slightly due to overlap of coordinate system
krig.AK$y.matrix[24]=2.4; krig.AK$y.meter[24]=2.4
krig.AK$x.matrix[24]=5.4; krig.AK$x.meter[24]=5.4

AK.map<-krig.AK %>% as.data.frame %>% 
  ggplot(aes(x.meter, y.meter, group=Sample)) + geom_point(aes(size=d15N, color=Sample), alpha=3/4) + 
  ggtitle("AK plot--d15N (permil)") + coord_equal() + theme_bw()

# just RK site
krig.RK<- scape.data %>% filter(Plot=="RK")
krig.RK$Sample<-droplevels(krig.RK$Sample)

RK.map<-krig.RK %>% as.data.frame %>% 
  ggplot(aes(x.meter, y.meter, group=Sample)) + geom_point(aes(size=d15N, color=Sample), alpha=3/4) +
  ggtitle("RK plot--d15N (permil)") + coord_equal() + theme_bw()

plot_grid(RK.map, AK.map, ncol=2, nrow=1)
dev.copy(pdf, "figures/dot.scape.pdf", width=9, height=4, encod="MacRoman")
dev.off()

#####################
#####################
```

AK d15N isoscape
```{r, krig AK d15N}
## Krig AK sites
coordinates(krig.AK)<- ~x.meter + y.meter

# Create a grid of "Pixels" using x as columns and y as rows
# And the rows and columns of pixels:
Columns=seq(from=-1, to=36, by=0.05)
Rows=seq(from=-1, to=21, by=0.05)
Grid.AK <- expand.grid(x=Columns,y=Rows)
coordinates(Grid.AK) <- ~ x+y
gridded(Grid.AK) <- TRUE # Plot the grid and points

## expand binding box 
#krig.AK@bbox # current data binding box
x<-c(-1, 36)
y<-c(-1, 21)
xy<-cbind(x,y)
S<-SpatialPoints(xy)
bbox(S)
krig.AK@bbox<-bbox(S) # expanded binding box for data
Grid.AK@bbox<-bbox(S) # expand for plot corner

# autokrige
d15N.AK.auto <- autoKrige(d15N ~ 1, krig.AK, Grid.AK) # ordinary kriging
plot(d15N.AK.auto, sp.layout = list(pts = list("sp.points", krig.AK)))
#d15N.AK.auto$krige_output

# make to dataframes for lm, combine
AK.pred.N <- d15N.AK.auto[1] %>% as.data.frame() # model predictions
AK.df.N <- krig.AK %>% as.data.frame() # dataframe

# make dataframe with data and predictions (krig) to assess fit
pred.AK.df.N<- left_join(AK.df.N, AK.pred.N, by = c("x.meter"="krige_output.x")) # longest axis is x
mod <- lm(krige_output.var1.pred ~ d15N, data=pred.AK.df.N)
summary(mod) #R squared = 0.21

mean(pred.AK.df.N$krige_output.var1.pred, na.rm=T) # mean of predictions 4.4 d15N
mean(pred.AK.df.N$krige_output.var1.stdev, na.rm=T) # 2.1

# inspect model
# plot(pred.AK.df.N$krige_output.var1.pred ~ pred.AK.df.N$d15N, xlab = "Observed", ylab = "Predicted", cex=0.2); abline(0,1, lty=2); abline(mod, col = "blue")

# Adding the sp.layout parameter shows the locations of the measurements
my.palette <- brewer.pal(n = 9, name = "RdBu")

### map in automap
#automapPlot(d15N.AK.auto$krige_output, "var1.pred", sp.layout = list("sp.points", krig.AK, pch=16, col="gray20", cex=0.5), col.regions=my.palette)

# map in ssplot
# note that the prediction "var1.pred" component of the krig is a SpatialPixelsDataFrame. Need to call the krig, then the component df("krige_output"), then the output column ("var1.pred") to plot in spplot

# variance
# spplot(d15N.AK.auto$krige_output,"var1.stdev")

#predicted krige
my.palette2 <- brewer.pal(n = 6, name = "BuGn")
my.palette3 <- colorRampPalette(c("firebrick1", "darkseagreen1", "azure1"))(4)
col.scheme.N <- colorRampPalette(my.palette3)


plotd15N.krig.AK<-spplot(d15N.AK.auto$krige_output["var1.pred"], col.regions=col.scheme.N, 
       sp.layout = list("sp.points", krig.AK, pch=c(16,17,1)[krig.AK$Sample], 
                        col=c("black"), cex=0.5, alpha=0.5),
       main=list(label=expression(paste(delta^{15}, N, " - AK (Restored Forest)")), cex=0.8), colorkey=FALSE)


###########################
```

RK d15N isoscape
```{r, krig RK d15N}
###########################
# RK Site Krig
###########################
## Krig RK site 
krig.RK<- scape.data %>% filter(Plot=="RK")
krig.RK$Sample<-droplevels(krig.RK$Sample)


coordinates(krig.RK)<- ~x.meter + y.meter

Columns=seq(from=-1, to=21, by=0.05)
Rows=seq(from=-1, to=36, by=0.05)
Grid.RK <- expand.grid(x=Columns,y=Rows)
coordinates(Grid.RK) <- ~ x+y
gridded(Grid.RK) <- TRUE # Plot the grid and points

## expand binding box 
#krig.RK@bbox # current data binding box
x<-c(-1, 21)
y<-c(-1, 36)
xy<-cbind(x,y)
S<-SpatialPoints(xy)
bbox(S)
Grid.RK@bbox<-bbox(S) # expand for plot corner

# autokrige
# 'fix.values' = The items describe the fixed value for the nugget (y-intercept), range (of interprolation) and sill(asymptote) respectively. Setting the value to NA means that the value is not fixed. Is passed on to autofitVariogram.
# The nugget is the y-intercept of the variogram. In practical terms, the nugget represents the small-scale variability of the data. A portion of that short range variability can be the result of measurement error. 
# The range is the distance after which the variogram levels off. The physical meaning of the range is that pairs of points that are this distance or greater apart are not spatially correlated. 
# The sill is the total variance contribution, or the maximum variability between pairs of points. 

# increasing the range gave better fit than auto
d15N.RK.auto <- autoKrige(d15N ~ 1, krig.RK, Grid.RK, fix.value=c(NA, 2, NA)) # ordinary kriging
plot(d15N.RK.auto, sp.layout = list(pts = list("sp.points", krig.RK)))
#d15N.RK.auto$krige_output

# make to dataframes for lm, combine
RK.pred.N <- d15N.RK.auto[1] %>% as.data.frame() # model predictions
RK.df.N <- krig.RK %>% as.data.frame() # dataframe

# make dataframe with data and predictions (krig) to assess fit
# longest axis here is y
pred.RK.df.N<- left_join(RK.df.N, RK.pred.N, by = c("x.meter"="krige_output.x"))
mod <- lm(krige_output.var1.pred ~ d15N, data=pred.RK.df.N)
summary(mod) #R squared = 0.21
mean(pred.RK.df.N$krige_output.var1.pred, na.rm=T) # mean of predictions 5.0 d15N
mean(pred.RK.df.N$krige_output.var1.stdev, na.rm=T) # 1.9

# inspect model
# plot(pred.RK.df.N$krige_output.var1.pred ~ pred.RK.df.N$d15N, xlab = "Observed", ylab = "Predicted", cex=0.2); abline(0,1, lty=2); abline(mod, col = "blue")

### map in automap
#automapPlot(d15N.RK.auto$krige_output, "var1.pred", sp.layout = list("sp.points", krig.RK), col.regions=my.palette)

# variance
# spplot(d15N.RK.auto$krige_output,"var1.stdev")

#predicted krige
plotd15N.krig.RK<-spplot(d15N.RK.auto$krige_output["var1.pred"], col.regions=col.scheme.N, 
       sp.layout = list("sp.points", krig.RK, pch=c(16,17,1)[krig.RK$Sample], 
                        col="black", cex=0.5, alpha=0.5), 
       main=list(label=expression(paste(delta^{15}, N, " - RK (Remnant Forest)")), cex=0.7))


#####################
grid.draw.ggmatrix <- function(x, recording = TRUE) {
  print(x)
}
### combined Ak-Rk krig d15N plot 
par(mfrow=c(1,2))
gridExtra::grid.arrange(plotd15N.krig.AK, plotd15N.krig.RK, ncol=2, nrow=1, widths = c(1.2,1))
dev.copy(png, "figures/RKAK.d15N.krig3.png", width = 8, height = 4, units = 'in', res = 300)
dev.off()


```


#### d13C isoscape  
setup
```{r, krig d13C}
################# d13C krigs
################# make site maps based on bubble plots

AK.map<-krig.AK %>% as.data.frame %>% 
  ggplot(aes(x.meter, y.meter, group=Sample)) + geom_point(aes(size=d13C, color=Sample), alpha=3/4) + 
  ggtitle("AK plot--d13C (permil)") + coord_equal() + theme_bw()


RK.map<-krig.RK %>% as.data.frame %>% 
  ggplot(aes(x.meter, y.meter, group=Sample)) + geom_point(aes(size=d13C, color=Sample), alpha=3/4) +
  ggtitle("RK plot--d13C (permil)") + coord_equal() + theme_bw()

plot_grid(RK.map, AK.map, ncol=2, nrow=1)
dev.copy(pdf, "figures/dot.scape.d13C.pdf", width=9, height=4, encod="MacRoman")
dev.off()

#####################
#####################
```

AK d13C isoscape
```{r, krig AK d13C}

# autokrige
d13C.AK.auto <- autoKrige(d13C ~ 1, krig.AK, Grid.AK) # ordinary kriging
plot(d13C.AK.auto, sp.layout = list(pts = list("sp.points", krig.AK)))
#d13C.AK.auto$krige_output

# make to dataframes for lm, combine
AK.pred.C <- d13C.AK.auto[1] %>% as.data.frame() # model predictions
AK.df.C <- krig.AK %>% as.data.frame() # dataframe

# make dataframe with data and predictions (krig) to assess fit
pred.AK.df.C<- left_join(AK.df.C, AK.pred.C, by = c("x.meter"="krige_output.x")) # longest axis is x
mod <- lm(krige_output.var1.pred ~ d13C, data=pred.AK.df.C)
summary(mod) #R squared = 0.24

mean(pred.AK.df.C$krige_output.var1.pred, na.rm=T) # mean of predictions -25.9 d13C
mean(pred.AK.df.C$krige_output.var1.stdev, na.rm=T) # 2.3

# inspect model
# plot(pred.AK.df.C$krige_output.var1.pred ~ pred.AK.df.C$d13C, xlab = "Observed", ylab = "Predicted", cex=0.2); abline(0,1, lty=2); abline(mod, col = "blue")

# Adding the sp.layout parameter shows the locations of the measurements
my.palette <- brewer.pal(n = 9, name = "RdBu")

# map in automap
#automapPlot(d13C.AK.auto$krige_output, "var1.pred", sp.layout = list("sp.points", krig.AK, pch=16, col="gray20", cex=0.5), col.regions=my.palette)

# map in ssplot
# variance
# spplot(d13C.AK.auto$krige_output,"var1.stdev")

#predicted krige
my.palette4 <- colorRampPalette(c("firebrick1", "paleturquoise2", "azure1"))(4)
col.scheme.C <- colorRampPalette(my.palette4)


plotd13C.krig.AK<-spplot(d13C.AK.auto$krige_output["var1.pred"], col.regions=col.scheme.C, 
       sp.layout = list("sp.points", krig.AK, pch=c(16,17,1)[krig.AK$Sample],
                        col="black", cex=0.5, alpha=0.5),
       main=list(label=expression(paste(delta^{13}, C, " - AK (Restored Forest)")), cex=0.8), colorkey=FALSE)


###########################
```

RK d13C isoscape
```{r, krig RK d13C}
###########################
# RK Site Krig
###########################
## Krig RK site 

# autokrige
d13C.RK.auto <- autoKrige(d13C ~ 1, krig.RK, Grid.RK, fix.value=c(NA, 1, NA)) # ordinary kriging
plot(d13C.RK.auto, sp.layout = list(pts = list("sp.points", krig.RK)))
#d13C.RK.auto$krige_output

# make to dataframes for lm, combine
RK.pred.C <- d13C.RK.auto[1] %>% as.data.frame() # model predictions
RK.df.C <- krig.RK %>% as.data.frame() # dataframe

# make dataframe with data and predictions (krig) to assess fit
# longest axis here is y
pred.RK.df.C<- left_join(RK.df.C, RK.pred.C, by = c("x.meter"="krige_output.x"))
mod <- lm(krige_output.var1.pred ~ d13C, data=pred.RK.df.C)
summary(mod) #R squared = 0.33

mean(pred.RK.df.C$krige_output.var1.pred, na.rm=T) # mean of predictions -27.6 d13C
mean(pred.RK.df.C$krige_output.var1.stdev, na.rm=T) # 2.8

# inspect model
# plot(pred.RK.df.C$krige_output.var1.pred ~ pred.RK.df.C$d13C, xlab = "Observed", ylab = "Predicted", cex=0.2); abline(0,1, lty=2); abline(mod, col = "blue")

# map in automap
#automapPlot(d13C.RK.auto$krige_output, "var1.pred", sp.layout = list("sp.points", krig.RK), col.regions=my.palette)

# variance
# spplot(d13C.RK.auto$krige_output,"var1.stdev")

#predicted krige
plotd13C.krig.RK<-spplot(d13C.RK.auto$krige_output["var1.pred"], col.regions=col.scheme.C, 
       sp.layout = list("sp.points", krig.RK, pch=c(16,17,1)[krig.RK$Sample],
                        col="black", cex=0.5, alpha=0.5), 
       main=list(label=expression(paste(delta^{13}, C, " - RK (Remnant Forest)")), cex=0.7))



#####################
grid.draw.ggmatrix <- function(x, recording = TRUE) {
  print(x)
}
### combined Ak-Rk krig d13C plot 
par(mfrow=c(1,2))
gridExtra::grid.arrange(plotd13C.krig.AK,plotd13C.krig.RK, ncol=2, nrow=1, widths = c(1.2,1))
dev.copy(png, "figures/RKAK.d13C.krig2.png", width = 8, height = 4, units = 'in', res = 300)
dev.off()


```

#### Density plots: all samples  
Density plots using predictions of N and C values for all sample types  
```{r, density plots fig S2, fig.show='hold', out.width='50%'}

###################
# make density plots using predictions
# combine the predictions for each plot
AK.df.out.N<-as.data.frame(AK.pred.N$krige_output.var1.pred); AK.df.out.N$Plot<-"AK"
RK.df.out.N<-as.data.frame(RK.pred.N$krige_output.var1.pred); RK.df.out.N$Plot<-"RK"; 
colnames(AK.df.out.N)<-c("d15N.pred", "Plot")
colnames(RK.df.out.N)<-c("d15N.pred", "Plot")

# new dataframe
pred.dat.N<-rbind(AK.df.out.N, RK.df.out.N)
pred.dat.N$Plot<-as.factor(pred.dat.N$Plot)
pred.dat.N %>% 
  group_by(Plot) %>%
  summarise(no_rows = length(Plot))

#man whitney for significance (not normal data)
mwu(pred.dat.N, d15N.pred, Plot) # signif difference in predictions (p<0.001)

# Density plot

#hist(pred.dat.N$d15N.pred[(pred.dat.N$Plot=="AK")], ylim=c(0,150000), xlab="d15N-predict")
#hist(pred.dat.N$d15N.pred[(pred.dat.N$Plot=="RK")], ylim=c(0,150000), xlab="d15N-predict")

# plot.means
predict.mean.N <- ddply(pred.dat.N, "Plot", summarise, grp.mean=mean(d15N.pred, na.rm=TRUE))

density.predict.N<-ggplot(pred.dat.N, aes(x=d15N.pred)) +
  geom_density(aes(color=Plot, fill=Plot), alpha=0.5)+
  xlab(expression(paste(delta^{15}, N["predicted"], sp="")))+
  scale_color_manual(values=c("#E69F00", "#56B4E9")) +
  scale_fill_manual(values=c("#E69F00", "#56B4E9")) +
  geom_vline(data=predict.mean.N, aes(xintercept=grp.mean, color=Plot),
             linetype="dashed", lwd=0.2) + BW.back2


######## Carbon


###################
# make density plots using predictions
# combine the predictions for each plot
AK.df.out.C<-as.data.frame(AK.pred.C$krige_output.var1.pred); AK.df.out.C$Plot<-"AK"
RK.df.out.C<-as.data.frame(RK.pred.C$krige_output.var1.pred); RK.df.out.C$Plot<-"RK"; 
colnames(AK.df.out.C)<-c("d13C.pred", "Plot")
colnames(RK.df.out.C)<-c("d13C.pred", "Plot")

# new dataframe
pred.dat.C<-rbind(AK.df.out.C, RK.df.out.C)
pred.dat.C$Plot<-as.factor(pred.dat.C$Plot)
pred.dat.C %>% 
  group_by(Plot) %>%
  summarise(no_rows = length(Plot))

#man whitney for significance (not normal data)
mwu(pred.dat.C, d13C.pred, Plot) # signif difference in predictions (p<0.001)

# Density plot

#hist(pred.dat.C$d13C.pred[(pred.dat.C$Plot=="AK")], ylim=c(0,150000), xlab="d13C-predict")
#hist(pred.dat.C$d13C.pred[(pred.dat.C$Plot=="RK")], ylim=c(0,150000), xlab="d13C-predict")

# plot.means
predict.mean.C <- ddply(pred.dat.C, "Plot", summarise, grp.mean=mean(d13C.pred, na.rm=TRUE))

density.predict.C<-ggplot(pred.dat.C, aes(x=d13C.pred)) +
  geom_density(aes(color=Plot, fill=Plot), alpha=0.5)+
  xlab(expression(paste(delta^{13}, C["predicted"], sp="")))+
  scale_color_manual(values=c("#E69F00", "#56B4E9")) +
  scale_fill_manual(values=c("#E69F00", "#56B4E9")) +
  geom_vline(data=predict.mean.C, aes(xintercept=grp.mean, color=Plot),
             linetype="dashed", lwd=0.2) + BW.back2

# get the legend, # create some space to the left of the legend
density.legend <- get_legend(
  density.predict.C + theme(legend.box.margin = margin(0, 0, 0, 12)) + 
    theme(legend.key.size = unit(0.3, "cm")))

gridExtra::grid.arrange(density.predict.N + theme(legend.position = "none"),
                        density.predict.C + theme(legend.position = "none"),
                        density.legend,
                        ncol=3, nrow=1, widths = c(1,1,0.4))

dev.copy(pdf, "figures/Fig S2ab.mod.pred.NC.pdf", width = 6, height = 2.5)
dev.off()
```
  
### Soil krigs    
  
#### d15N Soil isoscape
  
Subset data and only examine SOIL.     
- *Restored forest (AK)* and *Remnant Forest (RK)*
```{r, AK soil krig}
###### AK site d15N

## d15N SOIL plot
krig.AK<- scape.data %>% filter(Plot=="AK") # just AK
AK.soil<-  krig.AK[(krig.AK$Sample=="Soil"),] #just soil
AK.soil$Sample<-droplevels(AK.soil$Sample)

## Krig AK sites
coordinates(AK.soil)<- ~x.meter + y.meter

# Create a grid of "Pixels" using x as columns and y as rows
# And the rows and columns of pixels:
Columns=seq(from=-1, to=36, by=0.05)
Rows=seq(from=-1, to=21, by=0.05)
Grid.AK <- expand.grid(x=Columns,y=Rows)
coordinates(Grid.AK) <- ~ x+y
gridded(Grid.AK) <- TRUE # Plot the grid and points

## expand binding box 
#AK.soil@bbox # current data binding box
x<-c(-1, 36)
y<-c(-1, 21)
xy<-cbind(x,y)
S<-SpatialPoints(xy)
bbox(S)
AK.soil@bbox<-bbox(S) # expanded binding box for data
Grid.AK@bbox<-bbox(S) # expand for plot corner

# autokrige
d15N.AK.soil <- autoKrige(d15N ~ 1, AK.soil, Grid.AK) # ordinary kriging
plot(d15N.AK.soil, sp.layout = list(pts = list("sp.points", AK.soil)))
#d15N.AK.soil$krige_output

# make to dataframes for lm, combine
AK.soil.pred.N <- d15N.AK.soil[1] %>% as.data.frame() # model predictions
AK.soil.df.N <- AK.soil %>% as.data.frame() # dataframe

# make dataframe with data and predictions (krig) to assess fit
pred.AK.soil.df.N<- left_join(AK.soil.df.N, AK.soil.pred.N, by = c("x.meter"="krige_output.x"))#longest axis is x
mod <- lm(krige_output.var1.pred ~ d15N, data=pred.AK.soil.df.N)
summary(mod) #R squared = 0.015

mean(pred.AK.soil.df.N$krige_output.var1.pred, na.rm=T) # mean of predictions 6.0 d15N
mean(pred.AK.soil.df.N$krige_output.var1.stdev, na.rm=T) # 0.35

# inspect model
# plot(pred.AK.soil.df.N$krige_output.var1.pred ~ pred.AK.soil.df.N$d15N, xlab = "Observed", ylab = "Predicted", cex=0.2); abline(0,1, lty=2); abline(mod, col = "blue")

# Adding the sp.layout parameter shows the locations of the measurements
my.palette <- brewer.pal(n = 9, name = "RdBu")

# map in automap
#automapPlot(d15N.AK.soil$krige_output, "var1.pred", sp.layout = list("sp.points", AK.soil, pch=16, col="gray20", cex=0.5), col.regions=my.palette)

# map in ssplot
# variance
# spplot(d15N.AK.soil$krige_output,"var1.stdev")

#predicted krige
plotd15N.krig.AK.soil<-spplot(d15N.AK.soil$krige_output["var1.pred"], col.regions=col.scheme.N, 
       sp.layout = list("sp.points", AK.soil, pch=1, 
                        col=c("black"), cex=0.5, alpha=0.5),
       main=list(label=expression(paste(delta^{15}, N, " - AK (Restored Forest)")), cex=0.8), colorkey=FALSE)


###########################
# RK Site Krig
###########################

## d15N SOIL plot
krig.RK<- scape.data %>% filter(Plot=="RK") # just AK
RK.soil<-  krig.RK[(krig.RK$Sample=="Soil"),] #just soil
RK.soil$Sample<-droplevels(RK.soil$Sample)

## Krig RK sites
coordinates(RK.soil)<- ~x.meter + y.meter

# Create a grid of "Pixels" using x as columns and y as rows
# And the rows and columns of pixels:
Columns=seq(from=-1, to=21, by=0.05)
Rows=seq(from=-1, to=36, by=0.05)
Grid.RK <- expand.grid(x=Columns,y=Rows)
coordinates(Grid.RK) <- ~ x+y
gridded(Grid.RK) <- TRUE # Plot the grid and points

## expand binding box 
#RK.soil@bbox # current data binding box
x<-c(-1, 21)
y<-c(-1, 36)
xy<-cbind(x,y)
S<-SpatialPoints(xy)
bbox(S)
RK.soil@bbox<-bbox(S) # expanded binding box for data
Grid.RK@bbox<-bbox(S) # expand for plot corner

# autokrige
d15N.RK.soil <- autoKrige(d15N ~ 1, RK.soil, Grid.RK) # ordinary kriging
plot(d15N.RK.soil, sp.layout = list(pts = list("sp.points", RK.soil)))
#d15N.RK.soil$krige_output

# make to dataframes for lm, combine
RK.soil.pred.N <- d15N.RK.soil[1] %>% as.data.frame() # model predictions
RK.soil.df.N <- RK.soil %>% as.data.frame() # dataframe

# make dataframe with data and predictions (krig) to assess fit
pred.RK.soil.df.N<- left_join(RK.soil.df.N, RK.soil.pred.N, by = c("x.meter"="krige_output.x"))#longest axis is x
mod <- lm(krige_output.var1.pred ~ d15N, data=pred.RK.soil.df.N)
summary(mod) #R squared = 0.01

mean(pred.RK.soil.df.N$krige_output.var1.pred, na.rm=T) # mean of predictions 6.4 d15N
mean(pred.RK.soil.df.N$krige_output.var1.stdev, na.rm=T) # 0.8

# inspect model
# plot(pred.RK.soil.df.N$krige_output.var1.pred ~ pred.RK.soil.df.N$d15N, xlab = "Observed", ylab = "Predicted", cex=0.2); abline(0,1, lty=2); abline(mod, col = "blue")

# Adding the sp.layout parameter shows the locations of the measurements
my.palette <- brewer.pal(n = 9, name = "RdBu")

# map in automap
#automapPlot(d15N.RK.soil$krige_output, "var1.pred", sp.layout = list("sp.points", RK.soil, pch=16, col="gray20", cex=0.5), col.regions=my.palette)

# map in ssplot
# variance
# spplot(d15N.RK.soil$krige_output,"var1.stdev")

#predicted krige
plotd15N.krig.RK.soil<-spplot(d15N.RK.soil$krige_output["var1.pred"], col.regions=col.scheme.N, 
       sp.layout = list("sp.points", RK.soil, pch=1, 
                        col=c("black"), cex=0.5, alpha=0.5),
       main=list(label=expression(paste(delta^{15}, N, " - RK (Remnant Forest)")), cex=0.8), colorkey=TRUE)


#####################
grid.draw.ggmatrix <- function(x, recording = TRUE) {
  print(x)
}
### combined Ak-Rk krig d15N plot 
par(mfrow=c(1,2))
gridExtra::grid.arrange(plotd15N.krig.AK.soil, plotd15N.krig.RK.soil, ncol=2, nrow=1, widths = c(1.2,1))
dev.copy(png, "figures/RKAK.d15N.soils.png", width = 8, height = 4, units = 'in', res = 300)
dev.off()

```

#### d13C Soil isoscape
```{r}
###### AK site d13C

# autokrige
d13C.AK.soil <- autoKrige(d13C ~ 1, AK.soil, Grid.AK) # ordinary kriging
plot(d13C.AK.soil, sp.layout = list(pts = list("sp.points", AK.soil)))
#d13C.AK.soil$krige_output

# make to dataframes for lm, combine
AK.soil.pred.C <- d13C.AK.soil[1] %>% as.data.frame() # model predictions
AK.soil.df.C <- AK.soil %>% as.data.frame() # dataframe

# make dataframe with data and predictions (krig) to assess fit
pred.AK.soil.df.C<- left_join(AK.soil.df.C, AK.soil.pred.C, by = c("x.meter"="krige_output.x"))#longest axis is x
mod <- lm(krige_output.var1.pred ~ d13C, data=pred.AK.soil.df.C)
summary(mod) #R squared = 0.008

mean(pred.AK.soil.df.C$krige_output.var1.pred, na.rm=T) # mean of predictions -24.4 d13C
mean(pred.AK.soil.df.C$krige_output.var1.stdev, na.rm=T) # 0.6

# inspect model
# plot(pred.AK.soil.df.C$krige_output.var1.pred ~ pred.AK.soil.df.C$d13C, xlab = "Observed", ylab = "Predicted", cex=0.2); abline(0,1, lty=2); abline(mod, col = "blue")

# Adding the sp.layout parameter shows the locations of the measurements
my.palette <- brewer.pal(n = 9, name = "RdBu")

# map in automap
#automapPlot(d13C.AK.soil$krige_output, "var1.pred", sp.layout = list("sp.points", AK.soil, pch=16, col="gray20", cex=0.5), col.regions=my.palette)

# map in ssplot
# variance
# spplot(d13C.AK.soil$krige_output,"var1.stdev")

#predicted krige
plotd13C.krig.AK.soil<-spplot(d13C.AK.soil$krige_output["var1.pred"], col.regions=col.scheme.C, 
       sp.layout = list("sp.points", AK.soil, pch=1, 
                        col=c("black"), cex=0.5, alpha=0.5),
       main=list(label=expression(paste(delta^{13}, C, " - AK (Restored Forest)")), cex=0.8), colorkey=FALSE)



###########################
# RK Site Krig
###########################

# autokrige
d13C.RK.soil <- autoKrige(d13C ~ 1, RK.soil, Grid.RK) # ordinary kriging
plot(d13C.RK.soil, sp.layout = list(pts = list("sp.points", RK.soil)))
#d13C.RK.soil$krige_output

# make to dataframes for lm, combine
RK.soil.pred.C <- d13C.RK.soil[1] %>% as.data.frame() # model predictions
RK.soil.df.C <- RK.soil %>% as.data.frame() # dataframe

# make dataframe with data and predictions (krig) to assess fit
pred.RK.soil.df.C<- left_join(RK.soil.df.C, RK.soil.pred.C, by = c("x.meter"="krige_output.x"))#longest axis is x
mod <- lm(krige_output.var1.pred ~ d13C, data=pred.RK.soil.df.C)
summary(mod) #R squared = 0.16

mean(pred.RK.soil.df.C$krige_output.var1.pred, na.rm=T) # mean of predictions 4.4 d13C
mean(pred.RK.soil.df.C$krige_output.var1.stdev, na.rm=T) # 2.1

# inspect model
# plot(pred.RK.soil.df.C$krige_output.var1.pred ~ pred.RK.soil.df.C$d13C, xlab = "Observed", ylab = "Predicted", cex=0.2); abline(0,1, lty=2); abline(mod, col = "blue")

# Adding the sp.layout parameter shows the locations of the measurements
my.palette <- brewer.pal(n = 9, name = "RdBu")

# map in automap
#automapPlot(d13C.RK.soil$krige_output, "var1.pred", sp.layout = list("sp.points", RK.soil, pch=16, col="gray20", cex=0.5), col.regions=my.palette)

# map in ssplot
# variance
# spplot(d13C.RK.soil$krige_output,"var1.stdev")

#predicted krige
plotd13C.krig.RK.soil<-spplot(d13C.RK.soil$krige_output["var1.pred"], col.regions=col.scheme.C, 
       sp.layout = list("sp.points", RK.soil, pch=1, 
                        col=c("black"), cex=0.5, alpha=0.5),
       main=list(label=expression(paste(delta^{13}, C, " - RK (Remnant Forest)")), cex=0.8), colorkey=TRUE)


#####################
grid.draw.ggmatrix <- function(x, recording = TRUE) {
  print(x)
}
### combined Ak-Rk krig d13C plot 
par(mfrow=c(1,2))
gridExtra::grid.arrange(plotd13C.krig.AK.soil, plotd13C.krig.RK.soil, ncol=2, nrow=1, widths = c(1.2,1))
dev.copy(png, "figures/RKAK.d13C.soils.png", width = 8, height = 4, units = 'in', res = 300)
dev.off()

```
  
Density plots for soil isoscapes  
```{r, soil density plots, fig.show='hold', out.width='50%'}
# make density plots using predictions
# combine the predictions for each plot


############## Nitrogen plot
AK.df.soil.N<-as.data.frame(pred.AK.soil.df.N$krige_output.var1.pred); AK.df.soil.N$Plot<-"AK"
RK.df.soil.N<-as.data.frame(pred.RK.soil.df.N$krige_output.var1.pred); RK.df.soil.N$Plot<-"RK" 
colnames(AK.df.soil.N)<-c("d15N.pred", "Plot")
colnames(RK.df.soil.N)<-c("d15N.pred", "Plot")

# new dataframe
pred.soil.N<-rbind(AK.df.soil.N, RK.df.soil.N)
pred.soil.N$Plot<-as.factor(pred.soil.N$Plot)
pred.soil.N %>% 
  group_by(Plot) %>%
  summarise(no_rows = length(Plot))

#man whitney for significance (not normal data)
mwu(pred.soil.N, d15N.pred, Plot) # signif difference in predictions (p<0.001)

# Density plot

#hist(pred.soil.N$d15N.pred[(pred.soil.N$Plot=="AK")], ylim=c(0,10000), xlab="d15N-predict")
#hist(pred.soil.N$d15N.pred[(pred.soil.N$Plot=="RK")], ylim=c(0,10000), xlab="d15N-predict")

# plot.means
predict.mean.N <- ddply(pred.soil.N, "Plot", summarise, grp.mean=mean(d15N.pred, na.rm=TRUE))

density.predict.Nsoil<-ggplot(pred.soil.N, aes(x=d15N.pred)) +
  geom_density(aes(color=Plot, fill=Plot), alpha=0.5)+ ylim(0,1.2) +
  xlab(expression(paste(delta^{15}, N["predicted"], sp="")))+
  scale_color_manual(values=c("#E69F00", "#56B4E9")) +
  scale_fill_manual(values=c("#E69F00", "#56B4E9")) +
  geom_vline(data=predict.mean.N, aes(xintercept=grp.mean, color=Plot),
             linetype="dashed", lwd=0.2) + BW.back2


############## Carbon
AK.df.soil.C<-as.data.frame(pred.AK.soil.df.C$krige_output.var1.pred); AK.df.soil.C$Plot<-"AK"
RK.df.soil.C<-as.data.frame(pred.RK.soil.df.C$krige_output.var1.pred); RK.df.soil.C$Plot<-"RK" 
colnames(AK.df.soil.C)<-c("d13C.pred", "Plot")
colnames(RK.df.soil.C)<-c("d13C.pred", "Plot")

# new dataframe
pred.soil.C<-rbind(AK.df.soil.C, RK.df.soil.C)
pred.soil.C$Plot<-as.factor(pred.soil.C$Plot)
pred.soil.C %>% 
  group_by(Plot) %>%
  summarise(no_rows = length(Plot))

#man whitney for significance (not normal data)
mwu(pred.soil.C, d13C.pred, Plot) # signif difference in predictions (p<0.001)

# Density plot

#hist(pred.soil.C$d13C.pred[(pred.soil.C$Plot=="AK")], ylim=c(0,10000), xlab="d13C-predict")
#hist(pred.soil.C$d13C.pred[(pred.soil.C$Plot=="RK")], ylim=c(0,10000), xlab="d13C-predict")

# plot.means
predict.mean.C <- ddply(pred.soil.C, "Plot", summarise, grp.mean=mean(d13C.pred, na.rm=TRUE))

density.predict.Csoil<-ggplot(pred.soil.C, aes(x=d13C.pred)) +
  geom_density(aes(color=Plot, fill=Plot), alpha=0.5)+ ylim(0,0.8) +
  xlab(expression(paste(delta^{13}, C["predicted"], sp="")))+
  scale_color_manual(values=c("#E69F00", "#56B4E9")) +
  scale_fill_manual(values=c("#E69F00", "#56B4E9")) +
  geom_vline(data=predict.mean.C, aes(xintercept=grp.mean, color=Plot),
             linetype="dashed", lwd=0.2) + BW.back2

# get the legend, # create some space to the left of the legend
density.legend <- get_legend(
  density.predict.Nsoil + theme(legend.box.margin = margin(0, 0, 0, 12)) + 
    theme(legend.key.size = unit(0.3, "cm")))

gridExtra::grid.arrange(density.predict.Nsoil + theme(legend.position = "none"),
                        density.predict.Csoil + theme(legend.position = "none"),
                        density.legend,
                        ncol=3, nrow=1, widths = c(1,1,0.3))
dev.copy(pdf, "figures/Fig Sx.mod.pred.pdf", width = 5, height = 2.5)
dev.off()

```

```{r, AK plants krig, eval= FALSE}
#### d15N Plants isoscape
Subset data and only examine PLANTS     
- *Restored forest (AK)* and *Remnant Forest (RK)*
###### AK site d15N

## d15N plants plot
krig.AK<- scape.data %>% filter(Plot=="AK") # just AK
AK.plants<-  krig.AK[!(krig.AK$Sample=="Soil"),] #just plants
AK.plants$Sample<-droplevels(AK.plants$Sample)

## Krig AK sites
coordinates(AK.plants)<- ~x.meter + y.meter

# Create a grid of "Pixels" using x as columns and y as rows
# And the rows and columns of pixels:
Columns=seq(from=-1, to=36, by=0.05)
Rows=seq(from=-1, to=21, by=0.05)
Grid.AK <- expand.grid(x=Columns,y=Rows)
coordinates(Grid.AK) <- ~ x+y
gridded(Grid.AK) <- TRUE # Plot the grid and points

## expand binding box 
#AK.plants@bbox # current data binding box
x<-c(-1, 36)
y<-c(-1, 21)
xy<-cbind(x,y)
S<-SpatialPoints(xy)
bbox(S)
AK.plants@bbox<-bbox(S) # expanded binding box for data
Grid.AK@bbox<-bbox(S) # expand for plot corner

# autokrige
d15N.AK.plants <- autoKrige(d15N ~ 1, AK.plants, Grid.AK, fix.value=c(0.2, 3, 0.6)) # ordinary kriging
plot(d15N.AK.plants, sp.layout = list(pts = list("sp.points", AK.plants)))
#d15N.AK.plants$krige_output

# make to dataframes for lm, combine
AK.plants.pred.N <- d15N.AK.plants[1] %>% as.data.frame() # model predictions
AK.plants.df.N <- AK.plants %>% as.data.frame() # dataframe

# make dataframe with data and predictions (krig) to assess fit
pred.AK.plants.df.N<- left_join(AK.plants.df.N, AK.plants.pred.N, by = c("x.meter"="krige_output.x"))#longest axis is x
mod <- lm(krige_output.var1.pred ~ d15N, data=pred.AK.plants.df.N)
summary(mod) #R squared = 0.19

mean(pred.AK.plants.df.N$krige_output.var1.pred, na.rm=T) # mean of predictions 1.6 d15N
mean(pred.AK.plants.df.N$krige_output.var1.stdev, na.rm=T) # 0.7

# inspect model
# plot(pred.AK.plants.df.N$krige_output.var1.pred ~ pred.AK.plants.df.N$d15N, xlab = "Observed", ylab = "Predicted", cex=0.2); abline(0,1, lty=2); abline(mod, col = "blue")

# Adding the sp.layout parameter shows the locations of the measurements
my.palette <- brewer.pal(n = 9, name = "RdBu")

# map in automap
#automapPlot(d15N.AK.plants$krige_output, "var1.pred", sp.layout = list("sp.points", AK.plants, pch=16, col="gray20", cex=0.5), col.regions=my.palette)

# map in ssplot
# variance
# spplot(d15N.AK.plants$krige_output,"var1.stdev")

#predicted krige
plotd15N.krig.AK.plants<-spplot(d15N.AK.plants$krige_output["var1.pred"], col.regions=col.scheme.N, 
       sp.layout = list("sp.points", AK.plants, pch=c(16,17)[AK.plants$Sample],
                        col=c("black"), cex=0.5, alpha=0.5),
       main=list(label=expression(paste(delta^{15}, N, " - AK (Restored Forest)")), cex=0.8), colorkey=FALSE)

plotd15N.krig.AK.plants


###########################
# RK Site Krig
###########################

## d15N plants plot
krig.RK<- scape.data %>% filter(Plot=="RK") # just AK
RK.plants<-  krig.RK[!(krig.RK$Sample=="Soil"),] #just plants
RK.plants$Sample<-droplevels(RK.plants$Sample)

## Krig RK sites
coordinates(RK.plants)<- ~x.meter + y.meter

# Create a grid of "Pixels" using x as columns and y as rows
# And the rows and columns of pixels:
Columns=seq(from=-1, to=21, by=0.05)
Rows=seq(from=-1, to=36, by=0.05)
Grid.RK <- expand.grid(x=Columns,y=Rows)
coordinates(Grid.RK) <- ~ x+y
gridded(Grid.RK) <- TRUE # Plot the grid and points

## expand binding box 
#RK.plants@bbox # current data binding box
x<-c(-1, 21)
y<-c(-1, 36)
xy<-cbind(x,y)
S<-SpatialPoints(xy)
bbox(S)
RK.plants@bbox<-bbox(S) # expanded binding box for data
Grid.RK@bbox<-bbox(S) # expand for plot corner

# autokrige
d15N.RK.plants <- autoKrige(d15N ~ 1, RK.plants, Grid.RK) # ordinary kriging
plot(d15N.RK.plants, sp.layout = list(pts = list("sp.points", RK.plants)))
#d15N.RK.plants$krige_output

# make to dataframes for lm, combine
RK.plants.pred.N <- d15N.RK.plants[1] %>% as.data.frame() # model predictions
RK.plants.df.N <- RK.plants %>% as.data.frame() # dataframe

# make dataframe with data and predictions (krig) to assess fit
pred.RK.plants.df.N<- left_join(RK.plants.df.N, RK.plants.pred.N, by = c("x.meter"="krige_output.x"))#longest axis is x
mod <- lm(krige_output.var1.pred ~ d15N, data=pred.RK.plants.df.N)
summary(mod) #R squared = 0.15

mean(pred.RK.plants.df.N$krige_output.var1.pred, na.rm=T) # mean of predictions 2.8 d15N
mean(pred.RK.plants.df.N$krige_output.var1.stdev, na.rm=T) # 0.5

# inspect model
# plot(pred.RK.plants.df.N$krige_output.var1.pred ~ pred.RK.plants.df.N$d15N, xlab = "Observed", ylab = "Predicted", cex=0.2); abline(0,1, lty=2); abline(mod, col = "blue")

# Adding the sp.layout parameter shows the locations of the measurements
my.palette <- brewer.pal(n = 9, name = "RdBu")

# map in automap
#automapPlot(d15N.RK.plants$krige_output, "var1.pred", sp.layout = list("sp.points", RK.plants, pch=16, col="gray20", cex=0.5), col.regions=my.palette)

# map in ssplot
# variance
# spplot(d15N.RK.plants$krige_output,"var1.stdev")

#predicted krige
plotd15N.krig.RK.plants<-spplot(d15N.RK.plants$krige_output["var1.pred"], col.regions=col.scheme.N, 
       sp.layout = list("sp.points", RK.plants, pch=c(16,17)[RK.plants$Sample],
                        col=c("black"), cex=0.5, alpha=0.5),
       main=list(label=expression(paste(delta^{15}, N, " - RK (Remnant Forest)")), cex=0.8), colorkey=TRUE)
plotd15N.krig.RK.plants


#####################
grid.draw.ggmatrix <- function(x, recording = TRUE) {
  print(x)
}
### combined Ak-Rk krig d15N plot 
par(mfrow=c(1,2))
gridExtra::grid.arrange(plotd15N.krig.AK.plants, plotd15N.krig.RK.plants, ncol=2, nrow=1, widths = c(1.2,1))
dev.copy(png, "figures/RKAK.d15N.plants.png", width = 8, height = 4, units = 'in', res = 300)
dev.off()

```
```{r, eval=FALSE}
#####################
# Carbon kriging, leave alone for now.
#####################

## Krig AK site d13C
coordinates(krig.AK)<- ~x.meter + y.meter
col.scheme.C <- colorRampPalette(c('coral', 'cadetblue2', 'gray5'))(5)

Grid.AK <- spsample(krig.AK, type='regular', n=1e5)
gridded(Grid.AK) <- TRUE

## d13C plot
d13C.krig.AK <- krige(d13C ~ 1, krig.AK, Grid.AK)
plotd13C.krig.AK<-spplot(d13C.krig.AK["var1.pred"], col.regions=colorRampPalette(col.scheme.C), 
       main=expression(paste(delta^{13}, C, "--AK (Afforested Koa Plot)")))

## Krig RK site d13C
coordinates(krig.RK)<- ~x.meter + y.meter
GridRK <- spsample(krig.RK, type='regular', n=1e5)
gridded(GridRK) <- TRUE

## d13C plot
d13C.krig.RK <- krige(d13C ~ 1, krig.RK, GridRK)
plotd13C.krig.RK<-spplot(d13C.krig.RK["var1.pred"], col.regions=colorRampPalette(col.scheme.C), 
       main=expression(paste(delta^{13}, C, "--RK (Remnant Koa Plot)")))


```
```{r, kriging test, include=FALSE, eval=FALSE}
### Testing Kriging
################################################

# load data
krig<-read.csv("data/kriging.test.csv")

# simulate data
krig <- krig %>% 
  mutate(d15N.ex = as.numeric(rnorm(n=166, mean=2, sd=3)))

# just AK site
krig.AK<- krig %>% filter(site=="AK")

# example map with 
krig.AK %>% as.data.frame %>% 
  ggplot(aes(x.meter, y.meter)) + geom_point(aes(size=d15N.ex), color="forestgreen", alpha=3/4) + 
  ggtitle("d15N simulated (permil)") + coord_equal() + theme_bw()


# convert to spatial data, now a SPDF class object with 5 data slots
coordinates(krig.AK)<- ~x.meter + y.meter
str(krig.AK)
# data: contains all the variables associated with different spatial locations
# coords slot:  a matrix of all spatial locations with corresponding values
# bbox:  the bounding box or the 4 corners denoting spatial extent
# proj4string: projection inform. ie., what are the coordinates in?

# can coerce back to dataframe for plotting etc
# krig.AK<-krig.AK %>% as.data.frame %>% glimpse


##### VARIOGRAM, must be the SPDF object
# variogram function can take two arguments: 
# first: how one or more variables interact spatially
# second: SPDF object where those variables reside.


# no trend variogram
# example: variogram(log(zinc)~1, meuse)
vario.notrend <- variogram(d15N.ex~1, krig.AK)

# residual variogram w.r.t. a linear trend:
# example variogram(log(zinc)~x+y, meuse)
vario.resid <- variogram(d15N.ex~x.meter+y.meter, krig.AK) # calculates sample variogram values

# directional variogram:
# example: variogram(log(zinc)~x+y, meuse, alpha=c(0,45,90,135))
# example: variogram(log(zinc)~1, meuse, width=90, cutoff=1300)
vario.dir1 <- variogram(d15N.ex~x.meter+y.meter, krig.AK, alpha=c(0,45,90,135))
vario.dir2 <- variogram(d15N.ex~1, krig.AK, width=90, cutoff=1300)


######### fitting the variogram 
# first: variogram testing
# second: model with parameters to be fit to sample variogram
vario.fit <- fit.variogram(vario.resid, model=vgm(1, "Sph", 900, 1)) # fit model

plot(vario.resid, vario.fit) # plot the sample values, along with the fit model


##########################
# Kriging 

# need to make a grid to match where kriging/estimation is done

grid<-makegrid(krig.AK, n=1000)
colnames(grid)<-c('x.meter', 'y.meter')
grid<-as.data.frame(grid)

#where we have measurements
plot1 <- krig.AK %>% as.data.frame %>%
  ggplot(aes(x.meter, y.meter)) + geom_point(size=1) + coord_equal() + 
  ggtitle("Points with measurements")

# where we WANT measurements
plot2 <- grid %>% as.data.frame %>%
  ggplot(aes(x.meter, y.meter)) + geom_point(size=0.5) + coord_equal() + 
  ggtitle("Points at which to estimate")

#### make the plots
grid.arrange(plot1, plot2, ncol = 2)

# Once we have the prepared all of the above, we are now ready to krige. This can be done with the gstat::krige function, which usually takes four arguments:
# The model formula.
#  An SPDF of the spatial domain that has measurements.
#  An SPDF of the spatial domain to krige over.
#  A variogram model fitted to the data.
# Note that the second and third arguments have to be SPDF’s and cannot just be dataframes.


# make coordinates for grid (as we did for test data)
coordinates(grid)<- ~ x.meter + y.meter

# fit the model for the krig object (grid) and use model fit from corresponding model data above
test.krig<-krige(d15N.ex~1, krig.AK, grid, model=vario.fit)

test.krig %>% as.data.frame %>%
  ggplot(aes(x=x.meter, y=y.meter)) + geom_tile(aes(fill=var1.pred)) + coord_equal() +
  scale_fill_gradient(low = "yellow", high="red") +
  scale_x_continuous(labels=comma) + scale_y_continuous(labels=comma) +
  theme_bw()
```


